# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r echo=TRUE}
##setwd("./RepData_PeerAssessment1")
steps<-read.csv("./activity/activity.csv")
```
## What is mean total number of steps taken per day?
#### Get averages by day
```{r echo =TRUE}
stepdays<-tapply(steps$steps, steps$date, sum) 
hist(stepdays, main="Steps per day", xla="steps", yla="number of days")
```
 #### Mean and median:  
```{r echo=TRUE}
mean(stepdays, na.rm=T) 
median(stepdays, na.rm=T) 
``` 

## What is the average daily activity pattern?
### Plot time series: 
```{r echo=TRUE} 
stepmin<-aggregate(steps$steps, list(steps$interval), mean, na.rm=T) 
names(stepmin)<-c("interval", "steps") 
plot(stepmin$interval, stepmin$steps, type="l") 
``` 
### Get the interval with the most steps: 
```{r echo=TRUE}
stepmin[order(stepmin$steps),][288,]$interval 
``` 
## Imputing missing values
 
```{r echo=TRUE}
 length(which(is.na(steps$steps))) 
``` 
#note: R incorrectly thinks the HHMM time is an integer, but we can work with that instead of having to convert.   
#Filling the missing values
```{r echo=TRUE}
 steps$hour = floor(steps$interval/100) 
 steps$period=floor(steps$hour/3) 
 steps$period<-factor(steps$period) 
 levels(steps$period)<-c("0-2", "3-5", "6-8", "9-11", "12-14", "15-17", "18-20", "21-23") 
 mod<-lm(steps ~ period, data=steps) 
 mod
```
# Now imputing (the instructions say to create a new dataset, but I'm just adding a variable to the existing dataset, which lends itself to more useful comparisons): 
 
```{r echo=TRUE}
 steps$stepsi<-steps$steps 
 steps$stepsi[is.na(steps$steps)]<-predict(mod, newdata=steps[is.na(steps$steps),]) 
``` 

```{r echo=TRUE} 
 stepdaysi<-tapply(steps$stepsi, steps$date, sum, na.rm=T) 
 stepdaysi 
 hist(stepdaysi, main="Steps per day (with imputed data)", xla="steps", yla="number of days", col="#ff99ff") 
``` 
*mean (with imputed data)*: 
```{r echo=TRUE} 
 mean(stepdaysi, na.rm=T) 
``` 
*median  (with imputed data)*: 
```{r echo=TRUE}
 median(stepdaysi, na.rm=T) 
``` 



## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE} 
 steps$ddate<-as.character(steps$date) 
 steps$ddate<-as.Date(steps$ddate, format="%Y-%m-%d") 
 steps$weekday<-weekdays(steps$ddate) 
 steps$weekend<-F 
 steps$weekend[steps$weekday %in% c("Saturday", "Sunday")]<-T 
  
 
 stepmin.i.weekdays<-aggregate(steps$stepsi[!steps$weekend], list(steps$interval[!steps$weekend]), mean, na.rm=T) 
 stepmin.i.weekends<-aggregate(steps$stepsi[steps$weekend], list(steps$interval[steps$weekend]), mean, na.rm=T) 
 names(stepmin.i.weekdays)<-c("interval", "steps") 
 names(stepmin.i.weekends)<-c("interval", "steps") 
  
 
 par(mfrow = c(2,1), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0)) 
 plot(stepmin.i.weekends$interval, stepmin.i.weekends$steps, pch="", ylab="Steps", xlab="", main="weekend", type="l", ylim=c(0,220), col="blue") 
 plot(stepmin.i.weekdays$interval, stepmin.i.weekdays$steps, pch="", ylab="Steps", xlab="", main="weekday", type="l",  ylim=c(0,220), col="darkred") 
``` 

